---
title: "Untitled"
author: "JJW"
date: "2022-10-05"
output: html_document
---

```{r}
#load packages
library(tidyverse)
library(here)
library(magrittr)
library(visdat)
library(DataExplorer)
library(FactoMineR)
library(factoextra)
library(corrplot)
```


```{r}
#load data from literature screen

full_data<-read_csv(here("data", "raw_data", "full_data_filtered.csv"))




```

```{r}
#index dataframe to filter out extraneous variables, references etc

full_data_filter1<-full_data[,c(18:50)]

#check for initial missing data
vis_miss(full_data_filter1)

#index again

full_data_filter2<-full_data_filter1[,c(1:7,13:16,21:22)]


vis_miss(full_data_filter2)
```


```{r}

#add latin names and orders for host and spillover species

#add orders so can order network
all_hosts<-read.csv(here::here("data", "raw_data", "All_hosts.csv"))

all_hosts_H<-all_hosts[,(c(1:6))]

```
```{r}
colnames(all_hosts_H)[1]<-"Host Species"
```
```{r}
full_data_filter2$`Host Species`<-str_to_title(full_data_filter2$`Host Species`)
full_data_filter2$`Spillover Species`<-str_to_title(full_data_filter2$`Spillover Species`)
full_data_filter2$Order<-str_to_title(full_data_filter2$Order)
full_data_filter2$Zoonotic<-str_to_title(full_data_filter2$Zoonotic)
full_data_filter2$Virus_Name<-str_to_title(full_data_filter2$Virus_Name)
#trial1<-left_join(all_data,all_hosts_H)

#replace and with comma

full_data_filter2$`Host Species`%<>%str_replace(("(?i) and "),", ")
```


```{r}
#make column of adjusted virus names with just influenza a and rota virus a
full_data_filter2%<>%mutate(Virus_name_2=Virus_Name)

full_data_filter2$Virus_name_2[grep("Influenza A",full_data_filter2$Virus_name_2)]<- "Influenza A" 

full_data_filter2$Virus_name_2[grep("Rotavirus",full_data_filter2$Virus_name_2)]<- "Rotavirus A" 

full_data_filter2<-full_data_filter2[,c(1,14,2:13)]
```
all_data%<>%row_to_names(row_number = 1)

#species_data<-all_data[,c(8,9)]



#colnames(species_data)[1]<-"Host_species"
#colnames(species_data)[2]<-"Spillover_species"
```{r}
all_hosts_H$`Host Species`%<>%as.character()
#species_data$Spillover_species%<>%as.character()
```

```{r}
#split multiple common names into indeivudal rows
 all_hosts_H %<>%
       mutate(`Host Species` = strsplit(as.character(`Host Species`), ", ")) %>%
       unnest(`Host Species`)
all_hosts_S<-all_hosts_H
colnames(all_hosts_S)[1]<-"Spillover Species"
colnames(all_hosts_S)[2]<-"sHostNameFinal"
colnames(all_hosts_S)[3]<-"sOrder"
colnames(all_hosts_S)[4]<-"sFamily"
colnames(all_hosts_S)[5]<-"sGenus"
colnames(all_hosts_S)[6]<-"sSpecies"
```

```{r}
#unnest multiple host and spillover species into individual entries
full_data_filter2%<>%
  mutate(`Host Species`=strsplit(as.character(`Host Species`), ", ")) %>%
       unnest(`Host Species`)

full_data_filter2%<>%
  mutate(`Spillover Species`=strsplit(as.character(`Spillover Species`), ", ")) %>%
       unnest(`Spillover Species`)
```

```{r}
#wrangl host species names
full_data_filter2$`Host Species`[grep("oar", full_data_filter2$`Host Species`)] <- "Wild Boar" 
full_data_filter2$`Host Species`[grep("attle", full_data_filter2$`Host Species`)] <- "Cattle" 
full_data_filter2$`Host Species`[grep("ig", full_data_filter2$`Host Species`)] <- "Pigs" 
full_data_filter2$`Host Species`[grep("swine", full_data_filter2$`Host Species`)] <- "Pigs" 
full_data_filter2$`Host Species`[grep("African", full_data_filter2$`Host Species`)] <- "African Buffalo" 
full_data_filter2$`Host Species`[grep("ison", full_data_filter2$`Host Species`)] <- "American Bison" 
full_data_filter2$`Host Species`[grep("red deer", full_data_filter2$`Host Species`)] <- "Red Deer" 
full_data_filter2$`Host Species`[grep("eep", full_data_filter2$`Host Species`)] <- "Sheep" 
full_data_filter2$`Host Species`[grep("ampire", full_data_filter2$`Host Species`)] <- "Vampire Bat" 
full_data_filter2$`Host Species`[grep("abbit", full_data_filter2$`Host Species`)] <- "European Rabbit" 
full_data_filter2$`Host Species`[grep("armoset", full_data_filter2$`Host Species`)] <- "Common Marmoset"
full_data_filter2$`Host Species`[grep("accoon", full_data_filter2$`Host Species`)] <- "Northern Raccoon"
full_data_filter2$`Host Species`[grep("Civet", full_data_filter2$`Host Species`)] <- "Masked Palm Civet"
full_data_filter2$`Host Species`[grep("De Brazza's", full_data_filter2$`Host Species`)] <- "De Brazza's Monkey"
full_data_filter2$`Host Species`[grep("Sootey", full_data_filter2$`Host Species`)] <- "Sooty Mangabey"
full_data_filter2$`Host Species`[grep("Hamster", full_data_filter2$`Host Species`)] <- "Common Hamster" 
full_data_filter2$`Host Species`[grep("Dogs", full_data_filter2$`Host Species`)] <- "Domestic Dog" 
full_data_filter2$`Host Species`[grep("Gorilla", full_data_filter2$`Host Species`)] <- "Lowland Gorilla" 
full_data_filter2$`Host Species`[grep("quireel", full_data_filter2$`Host Species`)] <- "Variegated Squirrel" 
full_data_filter2$`Host Species`[grep("Red-Tailed", full_data_filter2$`Host Species`)] <- "Red-tailed Monkey" 
full_data_filter2$`Host Species`[grep("White-Toothed", full_data_filter2$`Host Species`)] <- "White-toothed Shrew" 
full_data_filter2$`Host Species`[grep("Crab-Eating", full_data_filter2$`Host Species`)] <- "Crab-eating Fox" 
```


```{r}
full_data_filter2%<>%
  mutate(hHostNameFinal2 = case_when(
    `Host Species`== "Goat" ~ "Capra_hircus", 
    `Host Species`== "Pigs" ~ "Sus_scrofa", 
   `Host Species`=="Cattle"~"Bos_taurus",
   `Host Species`=="Bovine"~"Bos_taurus",
   `Host Species`=="Dromedary Camel"~"Camelus_dromedarius",
   `Host Species`=="Sheep"~"Ovis_aries",
   `Host Species`=="Domestic Dog"~"Canis_familiaris",
   `Host Species`=="Domestic Cat"~"Felis_catus",
   `Host Species`=="Rat"~"Rattus_norvegicus",
   `Host Species`=="Poultry"~"Gallus_gallus",
   `Host Species`=="Chicken"~"Gallus_gallus",
   `Host Species`=="Turkey"~"Meleagris_gallopavo",
   `Host Species`=="Horse"~"Equus_caballus"))


```

```{r}
#repeat on spillover species
full_data_filter2$`Spillover Species`[grep("oar", full_data_filter2$`Spillover Species`)] <- "Wild Boar" 
full_data_filter2$`Spillover Species`[grep("attle", full_data_filter2$`Spillover Species`)] <- "Cattle" 
full_data_filter2$`Spillover Species`[grep("ig", full_data_filter2$`Spillover Species`)] <- "Pigs" 
full_data_filter2$`Spillover Species`[grep("swine", full_data_filter2$`Spillover Species`)] <- "Pigs" 
full_data_filter2$`Spillover Species`[grep("African", full_data_filter2$`Spillover Species`)] <- "African Buffalo" 
full_data_filter2$`Spillover Species`[grep("ison", full_data_filter2$`Spillover Species`)] <- "American Bison" 
full_data_filter2$`Spillover Species`[grep("red deer", full_data_filter2$`Spillover Species`)] <- "Red Deer" 
full_data_filter2$`Spillover Species`[grep("eep", full_data_filter2$`Spillover Species`)] <- "Sheep" 
full_data_filter2$`Spillover Species`[grep("ampire", full_data_filter2$`Spillover Species`)] <- "Vampire Bat" 
full_data_filter2$`Spillover Species`[grep("abbit", full_data_filter2$`Spillover Species`)] <- "European Rabbit" 
full_data_filter2$`Spillover Species`[grep("armoset", full_data_filter2$`Spillover Species`)] <- "Common Marmoset"
full_data_filter2$`Spillover Species`[grep("accoon", full_data_filter2$`Spillover Species`)] <- "Northern Raccoon"
full_data_filter2$`Spillover Species`[grep("Civet", full_data_filter2$`Spillover Species`)] <- "Masked Palm Civet"
full_data_filter2$`Spillover Species`[grep("De Brazza's", full_data_filter2$`Spillover Species`)] <- "De Brazza's Monkey"
full_data_filter2$`Spillover Species`[grep("Sootey", full_data_filter2$`Spillover Species`)] <- "Sooty Mangabey"
full_data_filter2$`Spillover Species`[grep("Hamster", full_data_filter2$`Spillover Species`)] <- "Common Hamster" 
full_data_filter2$`Spillover Species`[grep("Dogs", full_data_filter2$`Spillover Species`)] <- "Domestic Dog" 
full_data_filter2$`Spillover Species`[grep("Gorilla", full_data_filter2$`Spillover Species`)] <- "Lowland Gorilla" 
full_data_filter2$`Spillover Species`[grep("quireel", full_data_filter2$`Spillover Species`)] <- "Variegated Squirrel" 
full_data_filter2$`Spillover Species`[grep("Red-Tailed", full_data_filter2$`Spillover Species`)] <- "Red-tailed Monkey" 
full_data_filter2$`Spillover Species`[grep("White-Toothed", full_data_filter2$`Spillover Species`)] <- "White-toothed Shrew" 
full_data_filter2$`Spillover Species`[grep("Crab-Eating", full_data_filter2$`Spillover Species`)] <- "Crab-eating Fox" 
full_data_filter2$`Spillover Species`[grep("Black-Backed", full_data_filter2$`Spillover Species`)] <- "Black-backed Jackal" 
full_data_filter2$`Spillover Species`[grep("Spotted Deer", full_data_filter2$`Spillover Species`)] <- "Axis Deer" 
full_data_filter2$`Spillover Species`[grep("Skunk", full_data_filter2$`Spillover Species`)] <- "Striped Skunk" 
```

```{r}
full_data_filter2%<>%
  mutate(sHostNameFinal2 = case_when(
    `Spillover Species`== "Goat" ~ "Capra_hircus", 
    `Spillover Species`== "Pigs" ~ "Sus_scrofa", 
   `Spillover Species`=="Cattle"~"Bos_taurus",
   `Spillover Species`=="Bovine"~"Bos_taurus",
   `Spillover Species`=="Dromedary Camel"~"Camelus_dromedarius",
   `Spillover Species`=="Sheep"~"Ovis_aries",
   `Spillover Species`=="Domestic Dog"~"Canis_familiaris",
   `Spillover Species`=="Domestic Cat"~"Felis_catus",
   `Spillover Species`=="Rat"~"Rattus_norvegicus",
   `Spillover Species`=="Poultry"~"Gallus_gallus",
   `Spillover Species`=="Chicken"~"Gallus_gallus",
   `Spillover Species`=="Turkey"~"Meleagris_gallopavo",
   `Spillover Species`=="Horse"~"Equus_caballus",
   `Spillover Species`=="Guinea Fowl"~"Numida_meleagris",
   `Spillover Species`=="Giant Panda"~"Ailuropoda_melanoleuca",
    `Spillover Species`=="Florida Panther"~"Puma_concolor",
    `Spillover Species`=="Tshushima Cat"~"Prionailurus_bengalensis",
    `Spillover Species`=="White-Eared Opossum"~"Didelphis_albiventris",
   `Spillover Species`=="Northern Elephant Seal"~"Mirounga_angustirostris",
   `Spillover Species`=="Ferret"~"Mustela_furo",
   `Spillover Species`=="Donkey"~"Equus_asinus"))


```
```{r}

initial_join1<-left_join(full_data_filter2,all_hosts_H[,c(1,2)])
# #mergeinitial_join1<-
#   merge(full_data_filter2,all_hosts_H, by=c("hHostNameFinal", "Host Species"))

initial_join2<-left_join(initial_join1,all_hosts_S[,c(1,2)])


#coalesce

initial_join2%<>% mutate(hHostNameFinal = coalesce(hHostNameFinal,hHostNameFinal2))
initial_join2%<>% mutate(sHostNameFinal = coalesce(sHostNameFinal,sHostNameFinal2))


#make disntinct data frame of host latin names to avoid duplicates
Distinct_hHost<-all_hosts_H%>%distinct(hHostNameFinal, .keep_all = T)

Distinct_sHost<-all_hosts_S%>%distinct(sHostNameFinal, .keep_all = T)
#join orders etc
initial_join3<-left_join(initial_join2,Distinct_hHost[,c(2:4)], by="hHostNameFinal")

initial_join4<-left_join(initial_join3,Distinct_sHost[,c(2:4)],by="sHostNameFinal")

#index to remove extra name columns
initial_join4<-initial_join4[,c(1:5,7,17,19,20,8,18,21,22,6,9:14)]
```
```{r}
#use distinct to make a dataframe with only one entry for each virus

Distinct_virusDF<-full_data_filter2%>%distinct(Virus_name_2, .keep_all = T)


table(Distinct_virusDF$Order)

```

```{r}

# initial_join2$`Host Species`%<>%as.factor()
# #test2<-
# full_data_filter2%<>%
#   mutate(hHostNameFinal2 = case_when(
#     `Host Species`== "Goat" ~ "Capra_hircus", 
#     `Host Species`== "Pigs" ~ "Sus_scrofa", 
#    `Host Species`=="Cattle"~"Bos_taurus",
#    `Host Species`=="Bovine"~"Bos_taurus",
#    `Host Species`=="Dromedary Camel"~"Camelus_dromedarius",
#    `Host Species`=="Sheep"~"Ovis_aries",
#    `Host Species`=="Domestic Dog"~"Canis_familiaris",
#    `Host Species`=="Domestic Cat"~"Felis_catus",
#    `Host Species`=="Rat"~"Rattus_norvegicus",
#    `Host Species`=="Poultry"~"Gallus_gallus",
#    `Host Species`=="Chicken"~"Gallus_gallus",
#    `Host Species`=="Rabbit"~"Oryctolagus_cuniculus"))
# 
# test3<-merge(test2, all_hosts_H, by="hHostNameFinal")

#test1<-initial_join2%>%dplyr::mutate(hHostNameFinal=if_else(`Host Species`=="Pigs","Sus_scrofa",""))%>%dplyr::mutate(hHostNameFinal=if_else(`Host Species`=="Goat","Capra_hircus",hHostNameFinal)
                                                                                   
```









```{r}

 #mutate cell tropism into individual binary columns of cell type
data_wrangle1 <- initial_join4 %>%

  mutate(CT_macrophages=if_else(grepl("(?i)macro|(?i)mono|(?i)dendr",Cell_Tropism),"macrophages","not_macrophages"))%>%
  mutate(CT_respiratory_epithelium=if_else(grepl("(?i)resp|(?i)Bronchi",Cell_Tropism),"respiratory_epithelium","not_respiratory_epithelium"))%>%
  mutate(CT_endothelium=if_else(grepl("(?i)endo",Cell_Tropism),"endothelium","not_endothelium"))%>%
  mutate(CT_lymphocytes=if_else(grepl("(?i)lymph",Cell_Tropism),"lymphocytes","not_lymphocytes"))%>%
  mutate(CT_hepatocytes=if_else(grepl("(?i)hepato",Cell_Tropism),"hepatocytes","not_hepatocytes"))%>%
   mutate(CT_epithelial=if_else(grepl("(?i)Epithelial, lymphoid cells",Cell_Tropism),"epithelium","not_epithelium"))%>%
 # mutate(CT_keratinocytes=if_else(grepl("(?i)kerat",Cell_Tropism),"keratinocytes","not_keratinocytes"))%>%
  mutate(CT_intestinal_epithelium=if_else(grepl("(?i)intest",Cell_Tropism),"intestinal_epithelium","not_intestinal_epithelium"))%>%
  mutate(CT_pantropic=if_else(grepl("(?i)Pantropic",Cell_Tropism),"pantropic","not_pantropic"))%>%
 # mutate(CT_fibroblasts=if_else(grepl("(?i)fibro",Cell_Tropism),"fibroblasts","not_fibroblasts"))%>%
  mutate(CT_neurons=if_else(grepl("(?i)neur",Cell_Tropism),"neurons","not_neurons"))#%>%
 # mutate(CT_monocytes=if_else(grepl("(?i)mono",Cell_Tropism),"monocytes","not_monocytes"))%>%
  #mutate(CT_dendrocytes=if_else(grepl("(?i)dendr",Cell_Tropism),"dendrocytes","not_dendrocytes"))#%>%
 # mutate(CT_platelets=if_else(grepl("(?i)plat",Cell_Tropism),"platelets","not_platelets"))%>%
#  mutate(CT_RBCs=if_else(grepl("(?i)red",Cell_Tropism),"RBCs","not_RBCs"))

#put monocytes and dendrocytes with macrophages

```
```{r}
#repeat for organs affected and spread mechanism
data_wrangle2<-data_wrangle1%>%
  mutate(OA_respiratory=if_else(grepl("(?i)lung|(?i)resp|(?i)trachea|(?i)bronch|(?i)pulmo",`Systems/organs affected`),"respiratory_system","not_respiratory_system"))%>%
  mutate(OA_lymphoid=if_else(grepl("(?i)lymph|(?i)marrow|(?i)tonsil|(?i)spleen|(?i)immun",`Systems/organs affected`),"immune_system","not_immune_system"))%>%
  mutate(OA_gastrointestinal=if_else(grepl("(?i)gasto|GI|(?i)stomach|(?i)intest",`Systems/organs affected`),"gastrointestinal_system","not_gastrointestinal_system"))%>%
    mutate(OA_CNS=if_else(grepl("CNS|(?i)nerv|(?i)neur|(?i)ocular",`Systems/organs affected`),"nervous_system","not_nervous_system"))%>%
    mutate(OA_pancreas=if_else(grepl("(?i)pancreas",`Systems/organs affected`),"pancreas","not_pancreas"))%>%
  mutate(OA_renal=if_else(grepl("(?i)kidney",`Systems/organs affected`),"renal_system","not_renal_system"))%>%
    mutate(OA_heart=if_else(grepl("(?i)heart",`Systems/organs affected`),"heart","not_heart"))%>%
  mutate(OA_musculoskeletal=if_else(grepl("(?i)muscle|(?i)musculo",`Systems/organs affected`),"musculoskeletal_system","not_musculoskeletal_system"))%>%
    mutate(OA_endothelium=if_else(grepl("(?i)blood vessel wall|(?i)vascular",`Systems/organs affected`),"endothelium","not_endothelium"))%>%
    mutate(OA_integument=if_else(grepl("(?i)integument",`Systems/organs affected`),"integument","not_integument"))%>%
    mutate(OA_reproductive=if_else(grepl("(?i)placenta|(?i)reproductive|(?i)gonad|(?i)fetus",`Systems/organs affected`),"reproductive_system","not_reproductive_system"))%>%
   # mutate(OA_ocular=if_else(grepl("(?i)ocular",`Systems/organs affected`),"ocular","not_ocular"))%>%
   # mutate(OA_fetus=if_else(grepl("(?i)fetus",`Systems/organs affected`),"fetus","not_fetus"))%>%
  #mutate(OA_bone=if_else(grepl("(?i)bone",`Systems/organs affected`),"bone","not_bone"))%>%
    mutate(OA_thyroid=if_else(grepl("(?i)thyroid",`Systems/organs affected`),"thyroid","not_thyroid"))%>%
    mutate(OA_liver=if_else(grepl("(?i)liver",`Systems/organs affected`),"liver","not_liver"))
```
```{r}
#spread mechanism
data_wrangle3<-data_wrangle2%>%
  mutate(SM_lymphatic=if_else(grepl("(?i)lymph|(?i)macro",Spread_Mechanism),"lymphatic","not_lymphatic"))%>%
    mutate(SM_neuronal=if_else(grepl("(?i)synaptic|(?i)neuro",Spread_Mechanism),"neuronal","not_neuronal"))%>%
    mutate(SM_epithelial=if_else(grepl("(?i)epith|(?i)kerat",Spread_Mechanism),"epithelial","not_epithelial"))%>%
   # mutate(SM_endothelial=if_else(grepl("(?i)endo",Spread_Mechanism),"endothelial","not_endothelial"))%>%
  mutate(SM_gastrointestinal=if_else(grepl("(?i)intest",Spread_Mechanism),"gastrointestinal","not_gastrointestinal"))%>%
 mutate(SM_viremia=if_else(grepl("(?i)virem",Spread_Mechanism),"viremia","not_viremia"))
```

```{r}
#cell receptor
data_wrangle4<-data_wrangle3%>%
  mutate(CR_integrins=if_else(grepl("(?i)integrin",Cell_Receptor),"integrins","not_integrins"))%>%
  mutate(CR_HSPG=if_else(grepl("(?i)HSGP|HSPG|heparan",Cell_Receptor),"HSPG","not_HSPG"))
```

```{r}
#exploratory data analysis


create_report(data_wrangle4)

plot_bar(data_wrangle4)
```

```{r}


#make Df suitable for factominer

view(data_wrangle4)

facto_table1<-data_wrangle4[,c(3,14,21:49)]
facto_table1<-data.frame(lapply(facto_table1,factor))


summary(facto_table1)
#we can make the order related supplemtnary and make pathology active
res.mca = MCA(facto_table1,  quali.sup=c(1))

plot(res.mca)


plot.MCA(res.mca, invisible=c("var","quali.sup"), cex=0.7)
plot.MCA(res.mca, invisible=c("ind","quali.sup"), cex=0.7)
plot.MCA(res.mca, invisible=c("ind"))
plot.MCA(res.mca, invisible=c("ind", "var"))

#plot ellipses for zoontoic
zoonotic_ellipse<-plotellipses(res.mca,keepvar = 2)

ggsave(here::here("results", "zoonotic_ellipse.png"),plot = zoonotic_ellipse)

fviz_mca_biplot(res.mca, 
               repel = TRUE, # Avoid text overlapping (slow if many point)
               ggtheme = theme_minimal())

fviz_mca_var(res.mca, 
             repel = TRUE, # Avoid text overlapping (slow)
             ggtheme = theme_minimal())

grp <- as.factor(facto_table1[, "Zoonotic"])
zoonotic_ellipse_var<-fviz_mca_biplot(res.mca, repel = TRUE, col.var = "#E7B800",
 habillage = grp, addEllipses = TRUE, ellipse.level = 0.95, invisible = "quali.sup")

zoonotic_ellipse_var
ggsave(here::here("results", "zoonotic_ellipse_var.png"),plot = zoonotic_ellipse_var)


# Select the top 30 contributing individuals
# And the top 10 variables
fviz_mca_biplot(res.mca,
               select.ind = list(contrib = 30),
               select.var = list(contrib = 10))


#variable correlation to dimension

dimension_cor<-fviz_mca_var(res.mca, choice = "mca.cor",
             repel = TRUE)

ggsave(here::here("results", "dimension_cor.png"),plot = dimension_cor)

#screeplot to explain % of variance explained by each dimension
fviz_screeplot(res.mca, addlabels = TRUE, ylim = c(0, 45))

# Change the color by groups, add ellipses
#fviz_mca_biplot(res.mca, label="var", col.var ="red",
 #  habillage=grp, addEllipses=TRUE, ellipse.level=0.95)


dimdesc(res.mca)
#eigen values for each dimension
eigenvalues<-get_eigenvalue(res.mca)

fgh<-res.mca$var


var <- get_mca_var(res.mca)
var

# Coordinates
head(var$coord)
# Cos2: quality on the factore map
head(var$cos2)
# Contributions to the principal components
head(var$contrib)


corrplot(var$cos2, is.corr=FALSE)
fviz_cos2(res.mca, choice = "var", axes = 1:2)


# Contributions of rows to dimension 1
fviz_contrib(res.mca, choice = "var", axes = 1, top = 15)
# Contributions of rows to dimension 2
fviz_contrib(res.mca, choice = "var", axes = 2, top = 15)
```

###repeating with distinc t viruses DF

```{r}
#repeat the above analysis with the distinct virus DF

 #mutate cell tropism into individual binary columns of cell type
DV1 <- Distinct_virusDF %>%

  mutate(CT_macrophages=if_else(grepl("(?i)macro|(?i)mono|(?i)dendr",Cell_Tropism),"macrophages","not_macrophages"))%>%
  mutate(CT_respiratory_epithelium=if_else(grepl("(?i)resp|(?i)Bronchi",Cell_Tropism),"respiratory_epithelium","not_respiratory_epithelium"))%>%
  mutate(CT_endothelium=if_else(grepl("(?i)endo",Cell_Tropism),"endothelium","not_endothelium"))%>%
  mutate(CT_lymphocytes=if_else(grepl("(?i)lymph",Cell_Tropism),"lymphocytes","not_lymphocytes"))%>%
  mutate(CT_hepatocytes=if_else(grepl("(?i)hepato",Cell_Tropism),"hepatocytes","not_hepatocytes"))%>%
   mutate(CT_epithelial=if_else(grepl("(?i)Epithelial, lymphoid cells",Cell_Tropism),"epithelium","not_epithelium"))%>%
 # mutate(CT_keratinocytes=if_else(grepl("(?i)kerat",Cell_Tropism),"keratinocytes","not_keratinocytes"))%>%
  mutate(CT_intestinal_epithelium=if_else(grepl("(?i)intest",Cell_Tropism),"intestinal_epithelium","not_intestinal_epithelium"))%>%
  mutate(CT_pantropic=if_else(grepl("(?i)Pantropic",Cell_Tropism),"pantropic","not_pantropic"))%>%
 # mutate(CT_fibroblasts=if_else(grepl("(?i)fibro",Cell_Tropism),"fibroblasts","not_fibroblasts"))%>%
  mutate(CT_neurons=if_else(grepl("(?i)neur",Cell_Tropism),"neurons","not_neurons"))#%>%
 # mutate(CT_monocytes=if_else(grepl("(?i)mono",Cell_Tropism),"monocytes","not_monocytes"))%>%
  #mutate(CT_dendrocytes=if_else(grepl("(?i)dendr",Cell_Tropism),"dendrocytes","not_dendrocytes"))#%>%
 # mutate(CT_platelets=if_else(grepl("(?i)plat",Cell_Tropism),"platelets","not_platelets"))%>%
#  mutate(CT_RBCs=if_else(grepl("(?i)red",Cell_Tropism),"RBCs","not_RBCs"))

#put monocytes and dendrocytes with macrophages

```
```{r}
#repeat for organs affected and spread mechanism
DV2<-DV1%>%
  mutate(OA_respiratory=if_else(grepl("(?i)lung|(?i)resp|(?i)trachea|(?i)bronch|(?i)pulmo",`Systems/organs affected`),"respiratory_system","not_respiratory_system"))%>%
  mutate(OA_lymphoid=if_else(grepl("(?i)lymph|(?i)marrow|(?i)tonsil|(?i)spleen|(?i)immun",`Systems/organs affected`),"immune_system","not_immune_system"))%>%
  mutate(OA_gastrointestinal=if_else(grepl("(?i)gasto|GI|(?i)stomach|(?i)intest",`Systems/organs affected`),"gastrointestinal_system","not_gastrointestinal_system"))%>%
    mutate(OA_CNS=if_else(grepl("CNS|(?i)nerv|(?i)neur|(?i)ocular",`Systems/organs affected`),"nervous_system","not_nervous_system"))%>%
    mutate(OA_pancreas=if_else(grepl("(?i)pancreas",`Systems/organs affected`),"pancreas","not_pancreas"))%>%
  mutate(OA_renal=if_else(grepl("(?i)kidney",`Systems/organs affected`),"renal_system","not_renal_system"))%>%
    mutate(OA_heart=if_else(grepl("(?i)heart",`Systems/organs affected`),"heart","not_heart"))%>%
  mutate(OA_musculoskeletal=if_else(grepl("(?i)muscle|(?i)musculo",`Systems/organs affected`),"musculoskeletal_system","not_musculoskeletal_system"))%>%
    mutate(OA_endothelium=if_else(grepl("(?i)blood vessel wall|(?i)vascular",`Systems/organs affected`),"endothelium","not_endothelium"))%>%
    mutate(OA_integument=if_else(grepl("(?i)integument",`Systems/organs affected`),"integument","not_integument"))%>%
    mutate(OA_reproductive=if_else(grepl("(?i)placenta|(?i)reproductive|(?i)gonad|(?i)fetus",`Systems/organs affected`),"reproductive_system","not_reproductive_system"))%>%
   # mutate(OA_ocular=if_else(grepl("(?i)ocular",`Systems/organs affected`),"ocular","not_ocular"))%>%
   # mutate(OA_fetus=if_else(grepl("(?i)fetus",`Systems/organs affected`),"fetus","not_fetus"))%>%
  #mutate(OA_bone=if_else(grepl("(?i)bone",`Systems/organs affected`),"bone","not_bone"))%>%
    mutate(OA_thyroid=if_else(grepl("(?i)thyroid",`Systems/organs affected`),"thyroid","not_thyroid"))%>%
    mutate(OA_liver=if_else(grepl("(?i)liver",`Systems/organs affected`),"liver","not_liver"))
```
```{r}
#spread mechanism
DV3<-DV2%>%
  mutate(SM_lymphatic=if_else(grepl("(?i)lymph|(?i)macro",Spread_Mechanism),"lymphatic","not_lymphatic"))%>%
    mutate(SM_neuronal=if_else(grepl("(?i)synaptic|(?i)neuro",Spread_Mechanism),"neuronal","not_neuronal"))%>%
    mutate(SM_epithelial=if_else(grepl("(?i)epith|(?i)kerat",Spread_Mechanism),"epithelial","not_epithelial"))%>%
   # mutate(SM_endothelial=if_else(grepl("(?i)endo",Spread_Mechanism),"endothelial","not_endothelial"))%>%
  mutate(SM_gastrointestinal=if_else(grepl("(?i)intest",Spread_Mechanism),"gastrointestinal","not_gastrointestinal"))%>%
 mutate(SM_viremia=if_else(grepl("(?i)virem",Spread_Mechanism),"viremia","not_viremia"))
```

```{r}
#cell receptor
DV4<-DV3%>%
  mutate(CR_integrins=if_else(grepl("(?i)integrin",Cell_Receptor),"integrins","not_integrins"))%>%
  mutate(CR_HSPG=if_else(grepl("(?i)HSGP|HSPG|heparan",Cell_Receptor),"HSPG","not_HSPG"))

DV4<-DV4[,c(1:6,9:14,17:45 )]

#make anthro into zoonotic. can't say if anto/zoo without species for direction
DV4%<>%
  mutate(Zoonotic = case_when(
    Zoonotic== "Anthroponotic" ~ "Zoonotic",
    Zoonotic== "Zoonotic" ~ "Zoonotic",
    Zoonotic== "Not_zoonotic" ~ "Not_zoonotic")) 
```

```{r}
#exploratory data analysis


create_report(DV4)

plot_bar(DV4)
```

```{r}


#make Df suitable for factominer

view(DV4)


DVfacto_table1<-DV4[,c(3,6,13:41)]
DVfacto_table1<-data.frame(lapply(DVfacto_table1,factor))

summary(DVfacto_table1)

#we can make the order related supplemtnary and make pathology active
res.mcaDV = MCA(DVfacto_table1,  quali.sup=c(1))

plot.MCA(res.mcaDV)
plot.MCA(res.mcaDV, invisible=c("var","quali.sup"), cex=0.7)
plot.MCA(res.mcaDV, invisible=c("ind","quali.sup"), cex=0.7)
plot.MCA(res.mcaDV, invisible=c("ind","quali.sup"), cex=0.7,selectMod="contrib 20")
plot.MCA(res.mcaDV, invisible=c("ind","quali.sup"), cex=0.7,selectMod="cos2 10")
plot.MCA(res.mcaDV, invisible=c("ind"))
plot.MCA(res.mcaDV, invisible=c("ind", "var"))



plotellipses(res.mcaDV, keepvar = 2)
plotellipses(res.mcaDV, keepvar = 12:15)

dimdesc(res.mcaDV)


fviz_mca_biplot(res.mcaDV, 
               repel = TRUE, # Avoid text overlapping (slow if many point)
               ggtheme = theme_minimal())


fviz_mca_var(res.mcaDV, choice = "mca.cor", 
            repel = TRUE, # Avoid text overlapping (slow)
            ggtheme = theme_minimal())



fviz_mca_var(res.mcaDV, 
             repel = TRUE, # Avoid text overlapping (slow)
             ggtheme = theme_minimal())



# Color by cos2 values: quality on the factor map
fviz_mca_var(res.mcaDV, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())


fviz_mca_var(res.mcaDV, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # avoid text overlapping (slow)
             ggtheme = theme_minimal()
             )

```


